---
title: "Assignment 7"
author: "The Major Miners"
date: "5 November 2017"
output: pdf_document
---

```{r setup, include=FALSE, warning=FALSE,message=FALSE,echo=FALSE}
library(arules)
library(arulesViz)
library(data.table)
```
## Question 1
### Importing the dataset
```{r}
optdigits<-read.csv('optdigits.csv',header=TRUE)
```
### Part a
### Setting seed to 10 and performing k means clustering
```{r}
set.seed(10)
fit<-kmeans(optdigits[1:64],10,iter.max=200)
str(fit)
#Matrix that records the number of instances of digits in each cluster
#Rows denote the cluster number
#Columns denote the digits
#fit$cluster has the cluster that each row belongs to
k<-matrix(nrow=10,ncol=10,0)
for(i in 1:length(fit$cluster))
{
     k[fit$cluster[i],optdigits$digit[i]+1]<-k[fit$cluster[i],optdigits$digit[i]+1]+1
     #optdigits$digit[i]+1 as it is indexed from 1 and digits start from 0
}
#The digits are from 0-9
colnames(k)<-c(0:9) #c(c())?
d<-vector()
#Labelling each cluster with the digit which has the maximum number of instances in it
for(i in 1:nrow(k))
{
  d[i]<-which.max(k[i,])-1
}
rownames(k)<-d
print(rownames(k))
print(k)
```
### Part b
```{r}
#Part 2
#the first cluster has 113 1s and 97 9s. It's close.
#The cluster is cluster 1
count<-1
for(i in 1:length(fit$cluster))
{
  if(fit$cluster[i]==1)
  {
     d[count]<-i
     count<-count+1
  }
}
#New matrix containing only the rows that got clustered into cluster 1
newopt<-optdigits[d,]
#Hierarchical Clustering
clusters<-hclust(dist(newopt[1:64]))

plot(cutree(clusters, k = 2)) #can choose number of branches or cut height
memships = cutree(clusters, k = 2)
cluster1 = subset(newopt, memships==1)
cluster2 = subset(newopt, memships==2)
table(cluster1$digit)
table(cluster2$digit)
```
### Part c
```{r}
clusnum <- vector()
clusindex <- vector()
fit$centers[1,]
#fit$centers[1,] is the set of centers for the first cluster. There are 10 clusters.
#Load test data

test<-read.csv('optdigits_test.csv',header=TRUE)
for(i in 1:nrow(test)){
  distance = .Machine$integer.max
  for(j in 1:10){ #there are 10 clusters
    if(dist(rbind(test[i,2:ncol(test)], fit$centers[j,])) < distance){
      distance = dist(rbind(test[i,2:ncol(test)], fit$centers[j,]))
      clusnum[i] = rownames(k)[j]
      clusindex[i] = j
    }
  }
}
#clusnum refers to the digit that matches the input
#imagenumber is the index of the image in the test data
print(clusnum)
imagenumber = c(1:20)
result = data.frame(imagenumber, clusnum, clusindex)
print(result)
```
### Part d
```{r}
#Printing the number of data points present under each label
length(cluster1$digit) #139 numbers
length(cluster2$digit) #124 numbers

cluster1$clusternumber = seq(0,0,length = nrow(cluster1))
cluster2$clusternumber = seq(0,0,length = nrow(cluster2))
#cluster1 is mostly 9
#cluster2 is mostly 1
#add the cluster number and merge them
for(row in 1:nrow(cluster1)){
  cluster1[row,"clusternumber"] = 1;
}
for(row in 1:nrow(cluster2)){
  cluster2[row,"clusternumber"] = 2;
}

final = rbind(cluster1, cluster2)

#We observe that two images, the 4th and the 13th, were classified into cluster 1. They are the test dataset. We predict clusindex for them.
testdata = test[c(4,13),] #the ones classified to clusindex 1
traindata = test[-c(4,13),] #the ones that weren't
test_labels = clusindex[c(4,13)]
train_labels = clusindex[-c(4,13)]
library(class)
knnpredicted<-knn(traindata,testdata,cl = train_labels,k=7,prob=TRUE)
table(knnpredicted)
```


## Question 2
### Importing the dataset and modifying it to make it suitable for computation
```{r}
hwr<-read.csv('handwriting_recognition.csv',header=TRUE)
hwr<-hwr[rep(row.names(hwr),hwr$Freq),]
hwr<-hwr[,c(2:4)]
```
### Association rules with default settings
```{r}
default<-apriori(hwr,control=list(verbose=FALSE))
default_dt<-as.data.frame(data.table(lhs=labels(lhs(default)),rhs=labels(rhs(default)),quality(default)))
default_dt<-default_dt[,c(1:5)]
print(default_dt)
```
### Association rules for the remaining parts
```{r}
rules<-apriori(hwr,parameter = list(support=0.001, confidence=0.001),control=list(verbose=FALSE))
```
### Subquestion 1
###{Artist,Female}=> Recognized
```{r}
part1<-subset(rules, lhs %ain% c("Profession=Artist","Gender=Female") & rhs %ain% c("Recognition=Recognized"))
part1_dt<-as.data.frame(data.table(lhs=labels(lhs(part1)),rhs=labels(rhs(part1)),quality(part1)))
part1_dt<-part1_dt[,c(1:5)]
print(part1_dt)
```
### Subquestion 2
### {Engineer}=>Male
```{r}
part2<-subset(rules,lhs %ain% c("Profession=Engineer") & rhs %ain% c("Gender=Male"))
part2<-part2[1]
part2_dt<-as.data.frame(data.table(lhs=labels(lhs(part2)),rhs=labels(rhs(part2)),quality(part2)))
part2_dt<-part2_dt[,c(1:5)]
print(part2_dt)
```
### Subquestion 3
### {Actor,Recognized} => Female
```{r}
part3<-subset(rules,lhs %ain% c("Profession=Actor","Recognition=Recognized") & rhs %ain% c("Gender=Female"))
part3_dt<-as.data.frame(data.table(lhs=labels(lhs(part3)),rhs=labels(rhs(part3)),quality(part3)))
part3_dt<-part3_dt[,c(1:5)]
print(part3_dt)
```
### Subquestion 4
### {Doctor,Male} => Unrecognized
```{r}
part4<-subset(rules,lhs %ain% c("Profession=Doctor","Gender=Male") & rhs %ain% c("Recognition=Unrecognized"))
part4_dt<-as.data.frame(data.table(lhs=labels(lhs(part4)),rhs=labels(rhs(part4)),quality(part4)))
part4_dt<-part4_dt[,c(1:5)]
print(part4_dt)
```


